---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Dependencies 
```{r}
library(tidyverse)
```
## Loading and preprocessing the data
- Unzip the activity file if it is there.
- Load the data file into a data frame called `activity`

```{r}
unzip("activity.zip")
activity <- read_csv("activity.csv")
```
## Histogram of total number of steps taker per day

```{r}
daily_steps <- activity %>% 
    drop_na() %>%
    select(date, steps) %>% 
    group_by(date) %>%
    summarize(total_daily_steps = sum(steps, na.rm=TRUE), .groups = 'drop')
g <- ggplot(daily_steps, aes(date, total_daily_steps))
g + geom_col()
```

## What is mean total number of steps taken per day?
```{r}
daily_steps <- activity %>% 
    drop_na() %>%
    select(date, steps) %>% 
    group_by(date) %>%
    summarize(total_daily_steps = sum(steps, na.rm=TRUE), .groups = 'drop')
mean_total_steps_per_day <- mean(daily_steps$total_daily_steps, na.rm=TRUE)
mean_total_steps_per_day
```
## What is the median total number of steps taken per day?
```{r}
daily_steps <- activity %>% 
    drop_na() %>%
    select(date, steps) %>% 
    group_by(date) %>%
    summarize(total_daily_steps = sum(steps, na.rm=TRUE), .groups = 'drop')
median_total_steps_per_day <- median(daily_steps$total_daily_steps, na.rm=TRUE)
median_total_steps_per_day
```

## What is the average daily activity pattern?

```{r}

```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
